## **Introduction**

Our motivation on choosing the German Credit data is mainly because we think that this type of exercise will enhance our skills as business analysts for a proper decision making.

The German Credit data has data on 1000 past credit applicants, described by 30 variables. Each applicant is rated as "Good" or "Bad" credit (encoded as 1 and 0 respectively in the response variable). We want to obtain a model that may be used to determine if new applicants present a good or bad credit risk.

## **EDA**

```{r, echo = FALSE, message = FALSE, warning=FALSE}
source(here::here("script/setup.R"))
```

```{r, echo=FALSE, message=FALSE}
german_credit <- read.csv2(here::here("data/GermanCredit.csv"), header = TRUE)
 
german_credit <- german_credit[,-1] #to remove the first column which is not important
                                 #YES it is, because id the identifier of the customer

n<-dim(german_credit)[1] 
p<-dim(german_credit)[2]

numericvar <- c("DURATION", "AMOUNT", "INSTALL_RATE", "AGE", "NUM_CREDITS", "NUM_DEPENDENTS")
catevar <- c( "CHK_ACCT", "HISTORY", "JOB", "SAV_ACCT", "EMPLOYMENT", "PRESENT_RESIDENT")
binaryvar <- c("NEW_CAR", "USED_CAR", "FURNITURE", "RADIO.TV", "EDUCATION", "RETRAINING", "MALE_DIV", "MALE_SINGLE", "MALE_MAR_or_WID", "CO.APPLICANT", "GUARANTOR", "REAL_ESTATE", "PROP_UNKN_NONE", "OTHER_INSTALL", "RENT", "OWN_RES", "TELEPHONE", "FOREIGN")
```

### **Missing values**

```{r echo=FALSE, message=FALSE}

introduce(german_credit[,-p]) #describe that table is complete
plot_intro(german_credit[,-p])

```

```{r echo=FALSE, message=FALSE}
nrows<-nrow(german_credit)
n.missing<-rowSums(is.na(german_credit))
sum(n.missing > 0)/nrows
#0 meaning that there is no missing data

n.complete<-sum(complete.cases(german_credit))
n.complete/nrows
#1 meaning that all rows are complete and sum of it is equivalent to the number of row so when it's divided by the numbers of row. It returns 1

plot_missing(german_credit[,-p])
#showing that there is no missing value
```

### **Overall Summary**

We consider that the most important variables in the data would be...\<cannot make a decision now. We will get back to this after\>...

> **Findings during the Exploratory data analysis**

-   Response Variable: has 70% good applicants and 30% bad applicants.
-   Duration: the average duration of credit in months is 20.9 months whereas the first and third quartiles are 12 and 24 months.
-   History: something important to mention on the credit history is that there is 29% of critical accounts.
-   Variables considered as Purposes of credit:
    -   New Car: 23% of users
    -   Used Car: 10% of users
    -   Furniture: 18% of users
    -   Radio / TV: 28% of users
    -   Education: 5% of users
    -   Retraining: \~ 10% of users
    -   There is a 6% of users that we don't know the purpose of the credit
-   Gender and status of applicants:
    -   5% are male and divorced
    -   55% are male and single
    -   9% are male and married/a widower
    -   31% of the others such as women in any status

```{r echo=FALSE, message=FALSE}
eda_per_variable <- german_credit %>%
  mutate(CHK_ACCT = case_when(
    CHK_ACCT == 0 ~ "< 0DM",
    CHK_ACCT == 1 ~ "0 < -- < 200DM",
    CHK_ACCT == 2 ~ ">= 200DM",
    CHK_ACCT == 3 ~ "No checking account")) %>%
  mutate(HISTORY = case_when(
    HISTORY == 0 ~ "No credit data",
    HISTORY == 1 ~ "All credits at this bank paid back dully",
    HISTORY == 2 ~ "Existing credits paid",
    HISTORY == 3 ~ "No checking account",
    HISTORY == 4 ~ "Critical account")) %>%
  mutate(`NEW_CAR, Positive(Yes)` = case_when(
    NEW_CAR == 0 ~ "No",
    NEW_CAR == 1 ~ "Yes")) %>%
  mutate(`USED_CAR, Positive(Yes)` = case_when(
    USED_CAR == 0 ~ "No",
    USED_CAR == 1 ~ "Yes")) %>%
  mutate(`FURNITURE, Positive(Yes)` = case_when(
    FURNITURE == 0 ~ "No",
    FURNITURE == 1 ~ "Yes")) %>%
  mutate(`RADIO.TV, Positive(Yes)` = case_when(
    RADIO.TV == 0 ~ "No",
    RADIO.TV == 1 ~ "Yes")) %>%
  mutate(`EDUCATION, Positive(Yes)` = case_when(
    EDUCATION == 0 ~ "No",
    EDUCATION == 1 ~ "Yes",
    EDUCATION == -1 ~ "Binary as -1 (OBS #37)")) %>%
  mutate(`RETRAINING, Positive(Yes)` = case_when(
    RETRAINING == 0 ~ "No",
    RETRAINING == 1 ~ "Yes")) %>%
  mutate(SAV_ACCT = case_when(
    SAV_ACCT == 0 ~ "< 100DM",
    SAV_ACCT == 1 ~ "100 <= -- < 500DM",
    SAV_ACCT == 2 ~ "500 <= -- < 1000DM",
    SAV_ACCT == 3 ~ ">= 1000DM",
    SAV_ACCT == 4 ~ "Unknown/No savings account")) %>%
  mutate(EMPLOYMENT = case_when(
    EMPLOYMENT == 0 ~ "Unemployed",
    EMPLOYMENT == 1 ~ "< 1 year",
    EMPLOYMENT == 2 ~ "1 <= -- < 4 years",
    EMPLOYMENT == 3 ~ "1 <= -- < 7 years",
    EMPLOYMENT == 4 ~ ">= 7 years")) %>%
  mutate(`MALE_DIV, Positive(Yes)` = case_when(
    MALE_DIV == 0 ~ "No",
    MALE_DIV == 1 ~ "Yes")) %>%
  mutate(`MALE_SINGLE, Positive(Yes)` = case_when(
    MALE_SINGLE == 0 ~ "No",
    MALE_SINGLE == 1 ~ "Yes")) %>%
  mutate(`MALE_MAR_or_WID, Positive(Yes)` = case_when(
    MALE_MAR_or_WID == 0 ~ "No",
    MALE_MAR_or_WID == 1 ~ "Yes")) %>%
  mutate(`CO.APPLICANT, Positive(Yes)` = case_when(
    CO.APPLICANT == 0 ~ "No",
    CO.APPLICANT == 1 ~ "Yes")) %>%
  mutate(`GUARANTOR, Positive(Yes)` = case_when(
    GUARANTOR == 0 ~ "No",
    GUARANTOR == 1 ~ "Yes",
    GUARANTOR == 2 ~ "Binary as 2 (OBS #234)")) %>%
  mutate(PRESENT_RESIDENT = case_when(
    PRESENT_RESIDENT == 1 ~ "< 1 year",
    PRESENT_RESIDENT == 2 ~ "1 <= -- < 2 years",
    PRESENT_RESIDENT == 3 ~ "2 <= -- < 3 years",
    PRESENT_RESIDENT == 4 ~ ">= 4 years")) %>%
  mutate(`REAL_ESTATE, Positive(Yes)` = case_when(
    REAL_ESTATE == 0 ~ "No",
    REAL_ESTATE == 1 ~ "Yes")) %>%
  mutate(`PROP_UNKN_NONE, Positive(Yes)` = case_when(
    PROP_UNKN_NONE == 0 ~ "No",
    PROP_UNKN_NONE == 1 ~ "Yes")) %>%
  mutate(`OTHER_INSTALL, Positive(Yes)` = case_when(
    OTHER_INSTALL == 0 ~ "No",
    OTHER_INSTALL == 1 ~ "Yes"))%>%
  mutate(`RENT, Positive(Yes)` = case_when(
    RENT == 0 ~ "No",
    RENT == 1 ~ "Yes")) %>%
  mutate(`OWN_RES, Positive(Yes)` = case_when(
    OWN_RES == 0 ~ "No",
    OWN_RES == 1 ~ "Yes")) %>%
  mutate(JOB = case_when(
    JOB == 0 ~ "Unemployed/unskilled non-resident",
    JOB == 1 ~ "Unskilled - resident",
    JOB == 2 ~ "Skilled employee/official",
    JOB == 3 ~ "Management/self-employed/highly qualified employee/officcer",
    JOB == 4 ~ "Critical account")) %>%
  mutate(`TELEPHONE, Positive(Yes)` = case_when(
    TELEPHONE == 0 ~ "No",
    TELEPHONE == 1 ~ "Yes")) %>%
  mutate(`FOREIGN,, Positive(Yes)` = case_when(
    FOREIGN == 0 ~ "No",
    FOREIGN == 1 ~ "Yes"))

eda_per_variable$Customers <- "Customers"
eda_per_variable <- eda_per_variable %>%
  select(c(-NEW_CAR,-USED_CAR,-FURNITURE,-RADIO.TV,-EDUCATION,-RETRAINING,-MALE_DIV,
           -MALE_SINGLE,-MALE_MAR_or_WID,-CO.APPLICANT,-GUARANTOR,-REAL_ESTATE,
           -PROP_UNKN_NONE,-OTHER_INSTALL,-RENT,-OWN_RES,-TELEPHONE,-FOREIGN)) %>%
  relocate(RESPONSE, .after = last_col())

tbl_summary(
  eda_per_variable,
  type = all_continuous() ~ "continuous2",
  by = Customers, # split table by group
  missing = "no", # don't list missing data separately
  statistic = list(all_continuous() ~ c("{median}, ({p25}, {p75})",
                                        "{mean}, ({min}, {max})"),
                   all_categorical() ~ "{n}, ({p}%)")) %>%
  add_n() %>% # add column with total number of non-missing observations
  modify_header(label = "**Variables**") %>% # update the column header
  modify_caption("**German Credit Data Summary**") %>%
  bold_labels()
```

### **Histogram to visualize numerical and categorical variables**

Histogram is to summarize the distribution of the data set thus binary variables are not included in the chart which have only 0 and 1 values.

```{r echo=FALSE, message=FALSE}
plot_histogram(
  german_credit[,c(catevar, numericvar)],
  binary_as_factor = FALSE,
  geom_histogram_args = list(bins = 30L),
  scale_x = "continuous",
  title = "Histogram for numerical and categorical variables",
  ggtheme = theme_gray(),
  theme_config = list(),
  nrow = 4L,
  ncol = 4L,
  parallel = FALSE
)
```

-   AGE, AMOUNT and DURATION: positive skewness meaning that the bank may expect frequent young clients which borrow small loan with approximately 0-25 months credit duration and a few elderly clients which borrow large loan with longer credit duration.
-   The most frequent category for CHK_ACCT is no checking account and only a few applicants have more than 200 DM. Majority of applicants have less than 100 DM in saving account.
-   Most applicants are employed with the skilled employee category. The most highest frequency is in the 1-4 years category.
-   The most common installment rate is 4% of disposable income and the most common number of existing credits at this bank is 1. Also, the larger parts of applicants are individuals for whom liable to provide maintenance. In general, applicants have been in the current residence for more than 1 year.
-   Most of the applicants are paid the existing credits back duly however, the second most frequent group is the critical account.

### **Box plot to interpret numerical data**

We use box plot to visualize quartile, median, skewness and outliers of numerical data. Thus, in this section, we select only DURATION, AMOUNT, INSTALL_RATE, AGE, NUM_CREDITS and NUM_DEPENDENTS. However, we also include EMPLOYMENT and PRESENT_RESIDENT. Although both are categorical variables, the values represent the length of employment and residency. <br />

```{r echo=FALSE, message=FALSE}
plot_boxplot(german_credit[,c(numericvar, "RESPONSE","EMPLOYMENT", "PRESENT_RESIDENT")], by= "RESPONSE", binary_as_factor = FALSE, ncol = 2, title = "Side-by-side boxplots")
```

Regarding the 'good' applicants from box plot, they tend to have one of the following characteristics: shorter credit duration, longer employment duration, lower credit amount. Regarding the AGE variable, the age profiles of good applicants tend to be slightly higher. Also, we spot a possible error which is 125 years old. As for the rest variables, INSTALL_RATE, NUM_CREDITS, NUM_DEPENDENTS and PRESENT_RESIDENT, there are no significant differences between good and bad applicants.

> **Errors Found and Modification**

-   EDUCATION: is a binary variable (0/1) but we notice from the description table that minimum value is -1 (observation number 37).
-   GUARANTOR: is a binary variable (0/1) but we notice from the description table that maximum value is 2 (observation number 234).
-   AGE: we spotted 125 years (the observation number 537).
-   PRESENT_RESIDENT: is a categorical data and we spotted that it contains 0 to 3 in in the pdf data description; however, it contains 1 to 4 in the actual data. <br /> <br /> Therefore, we believe that these numbers are errors on the data, so we modify them from the `german_credit table` as following:
      -   EDUCATION at the observation number 37 is modified from -1 to 1
      -   GUARANTOR at the observation number 234 is modified from 2 to 1
      -   AGE at the observation 537 is modified from 125 to 75

```{r echo=FALSE, message=FALSE}
german_credit$EDUCATION[37] <- as.integer(1)
german_credit$GUARANTOR[234] <- as.integer(1)
german_credit$AGE[537] <- as.integer(75)

```

### **Summary by RESPONSE**

We select to show mean in this case which is more appropriate than median since our data contains many binary variables

```{r echo=FALSE, message=FALSE}
round(t(apply(german_credit[,c(numericvar, binaryvar)], 2, tapply, german_credit$RESPONSE, mean)), 2)
```

We can see that around 31% and 21% of total 'Good' credit used for radio/TV and new cars respectively while approximately 30% and 21% of 'Bad' credit used for new cars and radio/TV respectively.

### **Scatter plots and Correlations**

We can ignore correlations between categorical and binary variables. <br />

```{r echo=FALSE, message=FALSE}
german_credit_temp <- german_credit
german_credit_temp$RESPONSE <- as.factor(german_credit_temp$RESPONSE)

ggpairs(german_credit_temp[,c(numericvar, "RESPONSE")], ggplot2::aes(colour=RESPONSE))
```

<br /> From the result, we see that AMOUNT and DURATION are correlated.
